# Bank and financial institutions, ADBL; their status, operation and rural financing system

## Financial institutions

Financial sector refers to all the wholesale, retail, formal and informal institutions in an economy offering financial services to consumers, businesses and other financial institutions. In its broadest definition, it includes everything from banks, stock exchanges, and insurers, to credit unions, microfinance institutions and money lenders (DFID, 2004).

**What does the financial system do and how could it promote economic growth?**^[The Finance and Growth Nexus" available at \url{www.philadelphiafed.org/files/br/brjf00ak.pdf}]

To understand how the financial system might influence economic growth, we need to review the roles of the financial system in greater detail.

1. Financial system mobilizes savings. Since an individual saver may be unable or unwilling to completely fund a borrower, financial markets and institutions pool the savings of diverse households and make these funds available for lending. This activity reduces the transaction costs associated with external finance for both firms and households. By going directly to a financial institution, firms seeking to borrow avoid the costs of having to contact a diverse group of savers. Similarly, savers avoid the costs of evaluating every potential borrower by placing their funds with a financial institution. 
2. Financial system allocates savings by determining which borrowers obtain loans. Since financial institutions are specialists, they can determine worthwhile investment opportunities and judge the creditworthiness of borrowers at lower cost than the average small investor. 
3. Financial system is to reduce risk by spreading investors' savings across many different investment opportunities. Spreading savings diversifies risk for households and reduces their exposure to the uncertainty associated with individual projects. This reduction in risk encourages savings.
4. Financial system derives from its ability to generate liquidity. Some investments with potentially high returns involve projects that require long-term commitments of capital. However, some investors may unexpectedly need access to their savings. Fortunately, when the financial system pools the investments of many households, it allocates funds to both short- and long-term projects. Thus, investors obtain higher returns on their savings than they would if their investments were limited to short-term projects, but they still have access to their savings in unforeseen circumstances. Further, mixing investments in this way ensures that worthwhile long-term projects are funded. 
5. Financial system facilitates trade by extending credit and guaranteeing payments. For example, currency, demand deposits, and credit card accounts all allow individuals to exchange goods and services without having to resort to barter. Additionally, letters of credit help firms order the inputs for current production when they experience delays in payment for past sales.
6. Financial system also exerts corporate control and monitors managers. Entrepreneurs' or managers' information about the operation and outcome of their projects tends to be superior to information that outside creditors and shareholders have. Insiders' attempts to exploit this informational advantage by engaging in opportunistic behavior would tend to discourage savings. For example, managers might underreport their firms' profits to lenders and shareholders in order to raise their own earnings. To offset this information advantage, banks monitor borrowers, and equity markets allow shareholders to discipline managers by voting out poor management.

A modern financial system is important in the economy in order to

- pool and utilize financial resources,
- reduce costs and risks,
- expand and diversify opportunities,
- increase the allocative efficiency of resources and
- promote the productivity and facilitate the economic growth.

As a country's financial sector develops and becomes more efficient, it enables individuals and institutions to channel savings to more productive enterprises. This offers the prospect of significant growth in GDP. Hence, while a sound and healthy financial system contributes to the growth of the real sector on a sustainable basis, uncertainty and fragility in the financial system adversely affect the objectives of economic development in general and poverty reduction in particular, as witnessed by the experiences of the financial crises of the 1990s.

Moreover, from the point of view of central banks, developments in the financial sector have a special role that distinguishes them from other services. Financial developments influence the long term output and employment growth, they can affect the wide range of indicators and tools that central banks use to monitor and model the behavior of the economy.

The efficiency of a financial system typically depends on several key factors, including its degree of integration, market competition, level of development and innovation capacity. Indeed, there is a significant amount of empirical evidence derived from cross-country data showing that countries in which more credit to the private sector is extended typically experience stronger economic growth.

### Status and history^[https://www.nrb.org.np/contents/uploads/2020/11/Special_Publications-Nepalese_-Financial_-System_Growth_-and-_Challenges_July-2009.pdf]

In Nepal, the process of financial liberalization was actually initiated in the mid-1980s when the government gave a green signal for the entry of commercial banks in joint venture with foreign banks. With the need to modernize banking services through the transfer of technology and managerial skills, Nabil Bank Limited was established in 1984 as the first joint-venture bank in Nepal, which was subsequently followed by the establishments of Nepal Investment Bank Limited and Standard Chartered Bank Nepal Limited.

The process of financial liberalization gained momentum in 1987/88 when Nepal entered into a three-year Structural Adjustment Programme (SAP) with the IMF. The broad aim of this programme was to increase the role of market forces in the financial system. The NRB initiated regular auctions of treasury bill since 1988/89 in order to introduce greater flexibility in the interest rate structure. It also redesigned its rediscounting rates in the form of three windows: basic rate, selective rate and lender of the last resort rate in 1989.

For reducing the dependence of commercial banks on the NRB for short-term borrowing requirements, a call money market was set up during 1988/89. During the same period, a comprehensive study entitled "Commercial Bank Problem Analysis and Strategy Study (CBPASS)" was undertaken in order to improve the financial and organizational structure of the two state-owned commercial banks (Nepal Bank Limited [NBL] and Rastriya Banijya Bank [RBB]). Some of the suggestions incorporated in the CBPASS were implemented in 1990/91^[These included the recapitalization of the banks, repayment of government guaranteed loans and issue of government securities for provision for loan loss.].

Because of the satisfactory results of the SAP, Nepal entered into another three-year Enhanced Structural Adjustment Facility (ESAF) with the IMF in 1992/93. During that period, the core focus of financial liberalization was directed towards a) allowing market forces to play a greater role in the financial system, b) enhancing private sector participation in the development of the financial system, c) increasing competition and thereby efficiency of the financial system, d) improving and increasing increase financial instruments and e) developing the capital market.

Under the loan assistance of the World Bank and grant assistance of the Department for International Development (DFID) of the United Kingdom, the Financial Sector Technical Assistance (FSTA) Project has been implemented since 1998. The re-engineering of the NRB, restructuring of NBL and RBB and capacity building in the financial sector are the main components of the Project. Under FSTA, Project, some progress has been made in strengthening the regulatory and supervisory capabilities of the NRB. A new licensing policy in 2002 (last revised in 2007) was framed for opening a new commercial bank or financial institution^[This policy clearly spells out the following aspects: (a) minimum paid-up capital required, (b) share ownership structure, (c) depositing amount of share capital committed by the promoters, (d) special provisions on minimum paid-up capital for establishing bank/financial institution in economically backward regions of the country, (e) promoters' educational qualification, working experience and financial disciplinary aspects, (f) provisions on sale or transfer of the promoter shares, and (g) application processing time.].

An array of reform measures pertaining to financial sector legislation included the enactment of a new Nepal Rastra Bank Act, 2002 and Debt Recovery Act, 2002. With the enactment of the Nepal Rastra Bank Act, 2002, the central bank of the country has become an independent body in executing its policies for ensuring sound, efficient and healthy financial system through prudent regulatory and supervisory mechanism^[According to the NRB Act, 2002, the major objectives of NRB are as follows: a) formulation of monetary and foreign exchange policies in order to maintain the stability of price and BoP for sustainable economic development; b) promoting stability and liquidity required in banking and financial sector; c) developing secure, healthy and efficient payment system; d) regulating, inspecting, supervising and monitoring the banking and financial system; and e) promoting entire banking and financial system of Nepal and enhancing its public credibility. The Act also makes the NRB mandatory to make public the monetary policy report. For details, see NRB (2002).]. A Debt Recovery Tribunal (DRT) was set up under the Debt Recovery Act, 2002 with the aim of improving the legal and judicial processes of recovering problem loans of banks and financial institutions.

Although the Credit Information Bureau (CIB) was established in 1989 under the initiation of Nepal Rastra Bank, it was registered as a company in September 2004 and started its operation as a company from March 2005. The CIB is the prime organization in the country acting as the repository of credit information of the customers and commercial borrowers of all the banks and financial institutions. It collects the credit information from the banks and financial institutions and disseminates them on demand. Regulatory provisions on blacklisting the borrower and credit information were revised. 

To make the legal structure pertaining to banks and financial institutions up-dated, strong and convenient for the purpose of making bank and financial sector transactions healthy, secured and sound, the Secured Transactions Act and Insolvency Act were made effective from November 16, 2006 and November 20, 2006, respectively.

With a view to develop the NRB into a modern central bank, capable in maintaining financial sector stability, the second phase of financial sector reform is being focused on updating the information technology, supervisory capability and human resource management. Moreover, the accounting system of the NRB is being updated to comply with the requirements of international accounting standard and Nepal accounting standard.

In 2004, Bank and Financial Institution Ordinance was enacted and the new legal framework replaced different fragmented legal framework governing the operations of banks and financial institutions^[The Bank and Financial Institution Ordinance, 2004 has made a provision of categorizing deposit-taking financial institutions into four categories: A, B, C and D. With the enactment of a single or umbrella act governing the functions of deposit-taking financial institutions in Nepal, the issues related to the establishment, operation and exit of banks and financial institutions have been clearly spelt-out in the new legal framework. It should be noted that the Bank and Financial Institution Ordinance 2004 was converted into Bank and Financial Institution Act (BFIA) 2006 and made effective from November 3, 2006.]. In the context of promulgation of the Banks and Financial Institutions Act (BFIA), the existing prudential regulations and directions, which were separately issued for banks and financial institutions, have already been revised and integrated into a unified directive and came into implementation from July 16, 2005. 

To minimize the risk involved in the financial sector, the regulation regarding risk management has already been issued. To monitor such system by the NRB, the policy of implementing risk-based supervision is also being pursued. To develop risk management system, the private sector has been encouraged to establish a credit rating agency. 

In reference to the building up of appropriate infrastructure for the implementation of BASEL II Accord since 2007, necessary directives, policies and provisions for the banks and financial institutions have been formulated in compliance with the accord. Further, as per BASEL II Accord, the regulatory provision has been prepared complying with the Simplified Standardized Approach (SSA) for capital base and Basic Indicator Approach (BIA) for operational risk. 

The large amount of non-performing assets (NPA) has been a major challenge for financial sector development and its stability^[The overall size of NPLs in the Nepalese banking system has been estimated at about 8.98 percent as of mid-January 2008.]. As a measure to reduce the NPA level, the NRB is coordinating with the government in strengthening the capacity of Debt Recovery Tribunal and establishing an ‘Asset Management Company' (AMC).

To discourage the possible fraud and crimes in the financial system, the "Anti-Money Laundering Act 2008" and "Banking Crime and Punishment Act, 2008" have been enacted on January 28, 2008 and February 5, 2008 respectively. In this connection, the Financial Intelligence Unit has been established in the NRB and the related by laws and directives are being prepared. 

The profile of all commercial banks and financial institutions along with their major financial indicators has been prepared and posted on the web site of NRB to provide adequate information on financial conditions of commercial banks and financial institutions to the public. In addition, the directives issued to the licensed financial institutions by the NRB on the basis of on-site and off-site supervision has been placed on the website from 2006/07 for the use of the general public.

The NRB as a regulator and supervisor of the banking and financial sector has been trying to ensure a healthy, and efficient financial sector by improving regulation on par with international standard. As such, the capital adequacy, single borrower limit, loan loss provisioning, professionalism of the promoter, standardized capital base, total deregulation in interest rate, open licensing policy, phasing out the directed credit, etc., have been the major prudential regulatory reforms undertaken by the NRB. 

The NRB as a central bank has been contributing through the sequential financial sector liberalization, extending financial services to the people, who were previously beyond the reach of banking and financial services, and the process of deregulation to micro finance activities. Similarly, the micro-finance activities are encouraged through the establishment of microfinance development banks, micro-finance institutions and non-government organizations together with the contribution made by the NRB in the self-reliance fund (SRF).

Against this perspective, the Government unveiled the National Micro-Finance Policy on May 4, 2008 with the objective of enhancing the capacity for microfinance. As per the Policy, a separate body would be set up to regulate and supervise the microfinance. The aim of the Policy is to increase access of microfinance to backward and financially weak families and women for improving employment opportunities. National Microfinance Development Fund would be established to mobilize resources from local and international communities. The NRB would monitor and supervise the implementation of this Policy. 

The NRB has revised some of the regulatory norms that are not consistent with the market oriented economic policies. The provision relating to the interest rate spread has been abolished and the mandatory requirement of investing in priority sector for commercial banks has been phased-out. From 2007/08 onwards, priority sector lending requirement will not be mandatory for commercial banks. 

In the process of accession to the World Trade Organization, Nepal had made commitments in financial services under the General Agreement on Trade in Services (GATS) and the Annex on Financial services. But, all the commitments are conditional upon entry requirement, domestic laws, rules and regulations as well as the terms and conditions of the NRB and Insurance Board. Financial services can carried out in Nepal through a domestically incorporated company. Branches of foreign banks are to be given a green signal for wholesale banking as of January 1, 2010. However, only those financial institutions that obtain a rating of at least 'B' from prominent global credit rating agencies such as Moody's and Standard and Poor's will be permitted to seek commercial business in the country. Against this perspective, considering the beginning of banking transactions by foreign banks in Nepal and opening of branches by the Nepalese banks abroad, the bank has prepared a new draft of 'Banks and Financial Institutions Act' as per international standard and submitted to the Ministry of Finance. The draft contains special and separate provisions for the licensing and the operation of foreign bank branches.

Due to the various measures initiated by the NRB in the financial sector of the country and thus generating a hospitable environment for the entry of new financial institutions, many financial institutions have been established and competition has been boosted to some extent. Overall, as of mid-January 2021. Table \ref{tab:status-institutions-in-nepal} and \ref{tab:financial-sector-stability-2020} summarizes the categories and respective number of banks and financial institutions. Commercial banks, development banks, finance companies and cooperatives and NGOs licensed to undertake limited banking businesses fall under the regulatory and supervisory jurisdiction of the NRB. Given the relatively small and under-developed economic base of the country, Nepal has a reasonably diversified financial sector as evidenced by the number and variety of institutions that play an active role in the sector.

```{r disaggregated-status-institutions-in-nepal}
read_csv("./data/banks_class-A_commercial.csv", skip = 3, show_col_types = FALSE) %>% 
  knitr::kable(booktabs = TRUE, caption = "Class A commercial banks of Nepal (by mid-January, 2021); Source: \\url{https://www.nrb.org.np/contents/uploads/2021/02/List-of-BFIs-Poush-2077-_-English.pdf}", longtable = TRUE, align = "l") %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::column_spec(column = 1:6, width = c("1.5em", "14em", "5em", "8em", "4em", "10em"))

read_csv("./data/banks_class-B_development_banks.csv", skip = 3, show_col_types = FALSE) %>% 
  knitr::kable(booktabs = TRUE, caption = "Class B development banks of Nepal (by mid-January, 2021); Source: \\url{https://www.nrb.org.np/contents/uploads/2021/02/List-of-BFIs-Poush-2077-_-English.pdf}", longtable = TRUE, align = "l") %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::column_spec(column = 1:6, width = c("1.5em", "14em", "5em", "8em", "4em", "10em"))

read_csv("./data/banks_class-C_finance_companies.csv", skip = 3, show_col_types = FALSE) %>% 
  knitr::kable(booktabs = TRUE, caption = "Class B finance companies of Nepal (by mid-January, 2021); Source: \\url{https://www.nrb.org.np/contents/uploads/2021/02/List-of-BFIs-Poush-2077-_-English.pdf}", longtable = TRUE, align = "l") %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::column_spec(column = 1:6, width = c("1.5em", "14em", "5em", "8em", "4em", "10em"))

read_csv("./data/banks_class-D_microfinance.csv", skip = 3, show_col_types = FALSE) %>% 
  knitr::kable(booktabs = TRUE, caption = "Class D microfinance of Nepal (by mid-January, 2021); Source: \\url{https://www.nrb.org.np/contents/uploads/2021/02/List-of-BFIs-Poush-2077-_-English.pdf}", longtable = TRUE, align = "l") %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::column_spec(column = 1:6, width = c("1.5em", "14em", "5em", "8em", "4em", "10em"))

read_csv("./data/banks_class-other-institutions.csv", skip = 2, show_col_types = FALSE) %>% 
  knitr::kable(booktabs = TRUE, caption = "Other financial institutions of Nepal (by mid-January, 2021); Source: \\url{https://www.nrb.org.np/contents/uploads/2021/02/List-of-BFIs-Poush-2077-_-English.pdf}", longtable = TRUE, align = "l") %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::column_spec(column = 1:4, width = c("1.5em", "16em", "6em", "10em"))

read_csv("./data/banks_infrastructure_development_bank.csv", show_col_types = FALSE) %>% 
  knitr::kable(booktabs = TRUE, caption = "Infrastructure development bank of Nepal (by mid-January, 2021); Source: \\url{https://www.nrb.org.np/contents/uploads/2021/02/List-of-BFIs-Poush-2077-_-English.pdf}", longtable = TRUE, align = "l") %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 8) %>% 
  kableExtra::column_spec(column = 1:6, width = c("1.5em", "14em", "5em", "8em", "4em", "10em"))
```

Development banks and finance companies are not permitted to take demand deposits or undertake foreign currency business, but are otherwise treated similarly to commercial banks. A number of development banks were established with active participation of private sector after the Development Bank Act 1996 became effective with the aim of enhancing agriculture, industry and commerce through the provision of credit. Prior to this Act, there were only two development banks namely Agricultural Development Bank (which has been converted into commercial bank) and Nepal Industrial Development Corporation, both owned by the government. At present, development banks are operating under Bank and Financial Institutions Act 2006 possessing the status of "B" class financial institutions. NRB issues necessary directives on regular basis for the effective regulation of these banks.

The establishment of finance companies speeded up after the first amendment of the Finance Companies Act 1986 in 1992. At present, these companies operate under Bank and Financial Institutions Act, 2006 as "C" class financial institutions. Cooperatives, on the other hand, have been set up under Cooperatives Act 1992. The NRB has licensed a small number of these cooperatives to undertake limited financial transactions. These cooperatives operate under the directives issued by the NRB. There are also NRB-licensed NGOs to undertake limited banking transactions in accordance with the provision of the Financial Intermediation Related Institutions Act 1999.

The Employees Provident Fund (EPF), established on September 16, 1962 under the Employees Provident Fund Act 1962, manages the provident fund of the government employees, army, police, teachers, government corporations, and of some private companies. On the other hand, the Citizen Investment Trust (CIT) mobilizes the private and institutional savings, extends loans and advances, and works as an issue manager. The Ministry of Finance regulates the CIT which was set up in March 18, 1991. The Postal Savings Bank, on the other hand, established under the Postal Service Department, Government of Nepal came into operation in 1976. This service has played a role in inculcating the habit of saving on the part of people in lower level income groups.

The financial sector also encompasses areas such as capital market and insurance. The stock exchange, being the important player of the capital market, has a significant role in the Nepalese financial market. Nepal Government, under a program initiated to reform capital markets converted Securities Exchange Center into Nepal Stock Exchange Limited (NEPSE) in 1993^[Securities Exchange Center was established in 1976 with an objective of facilitating and promoting the growth of capital markets. Before conversion into stock exchange it was the only capital markets institution undertaking the job of brokering, underwriting, managing public issue, market making for government bonds and other financial services.]. The NEPSE's primary objective is to impart free marketability and liquidity to the government and corporate securities by facilitating transactions in its trading floor through member, market intermediaries, such as broker, market makers etc. NEPSE opened its trading floor on January 13, 1994. Secondary market transaction of Government bonds commenced at NEPSE on December 15, 2006. The NEPSE introduced fully automated screen based trading since August 24, 2007, replacing the previous open-out-cry system. It facilitates trading in the following instruments: a) shares (equity shares and preference shares), b) debentures, c) government bonds, and d) mutual funds. Further, process has begun to enter more securities brokers into the securities market and make it more competitive as per the provision in the Securities Broker Regulations, 2007. Moreover, the NEPSE has reconstructed its website and has started to include real time information since October 2007.

The Securities Act 2007 came into effect on January 14, 2007. The Act has further broadened the mandate of the Securities Board of Nepal (SEBON) as the regulator of the securities market. The SEBON has introduced Security Board Regulations 2007, Securities Market Operation Regulation 2007, and Securities Dealers Regulation 2007 on November 5, 2007 and Merchant Banker Regulation on March 15, 2008. It has developed Securities Registration and Issue Regulation, 2073.

The insurance sector was opened to the private sector in 1990. There are altogether 40 insurance companies established under Insurance Act, 1992 and Insurance Rules, 2049. 19 of these are life insurance, 20 non-life insurance, and 1 is re-insurance company^[https://nib.gov.np/wp-content/uploads/2020/05/Namelist-Insurance-Companies-2077.pdf]. The largely government-controlled National Insurance Corporation is the oldest corporate entity in this sector. The Insurance Board regulates the insurance sector.

### Challenges

- Establishment of banks, financial institutions and insurance companies help develop innovation and entrepreneurship that support long-term economic development. With the development of the financial sector, the challenge is to promote financial stability and discipline. Presently, the issue of ensuring effectiveness, transparency and discipline has also been of concern at the global level. Hence, it is also crucial to maintain financial stability in Nepal. The major challenge lies in enhancing competitiveness, expanding the financial access, and ensuring merger and acquisition besides strengthening systemic efficiency, discipline, and financial consolidation.

- Although recent measures undertaken by the NRB has improved the performance of the financial sector, the sector is still fragile and there is little access to financial services for the small businesses and low-income households. In this context, a recent study has identified two initiatives that the Government could take to help banks increase small business lending: a) develop a technical assistance fund to help potential banks build up suitable products and procedures for profitable lending to small businesses; b) create an enabling environment so that small business lending is safer, cheaper and faster. Similarly, for facilitating microfinance institutions serve a large number of low-income households, two steps could be undertaken by the government. The first one relates to promoting the microfinance industry by upgrading technical skills, reenergizing the sector, and reforming state-owned providers. The second measure concerns the creating of a legal and regulatory environment that protects microfinance consumers and promotes stability.

- With Nepal's membership to the World Trade Organization (WTO), the changes and reforms that the banks and financial institutions have to make are more intricate and challenging. The accession of Nepal to this global organization exemplifies both opportunities and risks in the process of maintaining overall financial stability so as to help accelerate the speed of economic growth. In the long run, it is indispensable for the banking sector to have the financial resources, skills and large-scale commitment to compete with larger competitors that have sophisticated system, capital to handle credit and operational risks.

- The financial services sector (FSS) of Nepal will be completely open for foreign participation after 2010 as per the commitments of FSS liberalization. However, experiences have demonstrated that liberalization coupled with inadequate strengthening of regulatory and supervisory framework, is an open invitation for financial crisis. Hence, the opportunities and threats of liberalization of the financial sector need to be properly accessed. In order to compete at par with foreign financial institutions the Nepalese financial institutions not only needs to upgrade its skills but also its size and reach. The ability to tap right technology becomes increasingly important to meet the more sophisticated demands. As such financial services entities will require large amounts of capital investments to remain competitive and be able to assume higher risks. 

- The country's financial sector has witnessed many changes in the past fostered, in part, by the emergence of new products and new market participants. While new financial products have undoubtedly led to better risk-sharing in the economy, they have also complicated the NRB's task of ensuring systemic stability through financial regulation and supervision. Moreover, due to a rapid increase in the number of banking and non-banking financial institutions with different modes of operations, the task of ensuring adequate monitoring and control by the NRB has been made more challenging.

- The development of credit rating agencies stands out as one of the possible options that would help reduce risk and check the high lending rates charged by financial intermediaries in the country. Equally important, credit rating agencies could help to de-emphasize the high importance attached by financial intermediaries to track records, a requirement that results from mistrust and lack of information about potential borrowers. The setting up of a credit rating agency is also necessary to provide information related to the securities and other market instruments besides contributing to the institutional good governance and practice of professional ratings of the institutions.

- An array of hindrances has restrained capital market development.A central depository system and a scriptless trading platform are prerequisites for an efficient capital market system but have been held up by the lack of an adequate legal framework. Current disclosure levels for listed companies are also weak and need to be supported by the development of a comprehensive set of accounting standards by the Institute of Chartered Accountants of Nepal. The nascent state of the insurance and pension sectors, and more generally, a limited investor and issuer base are also impediments to the development of the private securities markets.

- Given the present context of lacking the appropriate investment infrastructure for institutional investors, such as the CIT, EPF, and insurance companies, there is a need to promote and manage the issuance of debentures, mutual funds, and unit plans through the stock exchange. Furthermore, to make the capital mobilization activities more professional and transparent, there is a need of clear regulatory measures for these institutions.

- Since the NEPSE is being run under the government ownership and management, it would be difficult for an autonomous regulatory institution to regulate and supervise it. In this context, there is a need to review the ownership and management and develop it as a modern institution operating as per the international practice.

- Given the growing concept of international listing and trading in the South Asian region, it is necessary to strengthen the professional and institutional capability of securities market and brokers. Establishment of a Central Depository System (CDS) is necessary in this regard. This will spread up the securities transfer process. Further, there is a challenge to develop the securities market, to make it a better avenue for NRNs and FIIs.

- Closer links between formal and informal financial markets should be developed by encouraging formal financial institutions to mobilize deposits and allocate credit through informal and community-based banks and microfinance agents in areas where the reach of formal banks is limited. Fiscal policies, as well as regulatory and supervisory structures, should be designed to encourage these developments.

- Capital account liberalization can play a useful role in financial reforms. Opening up to foreign banks and other financial firms and to foreign direct investment in the financial sector has many potential benefits. These benefits include the introduction of financial innovations and sophisticated financial instruments by foreign financial firms, added depth in domestic financial markets due to foreign inflows, and more efficiency in the domestic banking sector through increased competition. The academic literature indicates, however, that precipitous opening of the capital account before the domestic financial sector has reached a certain level of maturity and the appropriate regulatory expertise is in place could spell trouble.

- Action should be taken to build local capabilities in risk management. This encompasses development of trust and policy credibility, as well as governance procedures. Macroeconomic risks can be reduced by pursuing sound macroeconomic policies, improving coordination between fiscal and monetary policies, and careful management of government borrowing. Market risks can be mitigated by improving capital market efficiency, reducing interest rate volatility, developing secondary markets for treasury bills and improving liquidity management. Microeconomic risks can be lowered by improving the accuracy, reliability and timeliness of financial information, enforcing financial contracts, providing efficient and reliable payment systems, risk-sharing and credit risk insurance schemes, and enhancing diversification in small markets. 

- An essential prerequisite of fiscal and monetary control is reliable and timely information on the financial and nonfinancial sectors. Decision-makers in central banks, supervisory agencies, and finance ministries need adequate financial information to monitor and diagnose financial sector problems and to distinguish between illiquidity and insolvency of individual institutions. Similarly, banks and investors need good financial information to assess and monitor the creditworthiness of borrowers and issuers of shares.

- Strong and reliable legal and regulatory frameworks are the foundations upon which policy credibility and financial stability rest. In their absence, economic and financial policies can not be implemented. Legal and regulatory enforcement is needed to ensure action is taken against those that violate their prudential and financial obligations, and to act as a credible deterrent for all market participants.

## Status of financial sector (Economic Survey, 2019-20)

The financial sector has been expanding rapidly in recent years. Access to banking, insurance and capital markets is increasing. The financial system has become more stable and financial instruments have become more productive. Investment is encouraged due to reforms in interest rates on loans. Financial services are being rural oriented and based on information technology. The share of bad loans of financial institutions is the lowest in South Asia. The scope of capital market is also widening with the use of electronic system in transactions.

Due to the effective implementation and improvement of the policy arrangements for the development and consolidation of the financial sector in the last few years, access to finance has been increasing and financial transactions have also expanded significantly.

```{r status-institutions-in-nepal}

tribble(
  ~"institution", ~"number", 
  "Banks and financial institutions", 162,
  "Insurance companies", 40,
  "Cooperatives", 34837,
  "Employees' provident fund", 1,
  "Citizen investment trust", 1,
  "Postal saving bank", 1
) %>% 
  knitr::kable(caption = "Institution as of mid-March 2020", booktabs = TRUE) %>% 
  kableExtra::kable_styling(latex_options = "striped")
```

The banking sector's share has been increasing in the asset/liability structure of Nepal's financial system. By the mid-January 2019, the share of bank and financial institutions to the total assets/liability structure of the financial system is 88.1 percent. In this structure, share of Nepal Rastra Bank is 16.8 percent and that of commercial banks is 57.6 percent. Thus, the ratio of total assets/liabilities of organizations operating in banking sector and contractual savings to GDP is 161.3 percent.

### Financial sector stability

The indicators of financial sector stability in fiscal year 2019/20 are robust in totality. But the ratio of non-performing loan (NPL) of banks and financial institutions has increased to 1.8 percent in mid-January 2020 from 1.5 percent of mid-July 2019.

```{r financial-sector-stability-2020}
read_csv("./data/major_indicators_financial_sector_stability.csv", skip = 2, show_col_types = FALSE) %>% 
  knitr::kable(caption = "Major indicators of financial sector stability (Source: NRB, 2020); by mid-January.", booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(font_size = 8) %>% 
  kableExtra::column_spec(column = 1:7, width = c("8em", rep("5em", 6)))
```

### Province-wise branches of bank and financial institutions

```{r provincewise-bfi-branches-2020}
read_csv("./data/province-wise_branches_of_bfi.csv", skip = 3, show_col_types = FALSE) %>% 
  knitr::kable(caption = "Province-wise branches of bank and financial institutions (Source: NRB, 2020; by mid-March 2020)", booktabs = TRUE) %>% 
  kableExtra::kable_styling(font_size = 8) %>% 
  kableExtra::column_spec(1:9, width = c("6em", rep("5em", 7)))
```

### Bank rate, re-financing rate and compulsory reserve ratio (in percent)

```{r bfi-refinancing-rate-2020}
read_csv("./data/bank_rate_refinancing_rate_compulsory_reserve_ratio.csv", skip = 3, show_col_types = FALSE) %>% 
  knitr::kable(caption = "Bank rate, re-financing rate and compulsory reserve ratio (in percent) (Source: NRB, 2020", booktabs = TRUE, longtable = TRUE) %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = "striped")
```

## Current status of national debt of Nepal

(Refer to Comprehensive Guide (Contemporary and GA) for a summary discussion of the topic.)

## Agriculture Development Bank Limited

### Establishment, vision, mission

### Contribution to agriculture promotion, rural poverty and economic growth

## Credit and finance management system

- CPAS software
- Credit administration (CAD) related services
