# Agriculture supply chain, project management, agribusiness and market

## Agriculture supply chain

## Project management

### Monitoring

Monitoring is the process of _keeping track of progress_ _on a continuous and/or periodic basis_ by management at different levels of an institutional hierarchy, or the individual or agency entrusted by the management to scrutinize whether the inputs and resources meant for the implementation of plans, policies, programmes and projects are being properly delivered. Furthermore, the role of monitoring is to verify whether the project activities are being implemented and whether or not the intended outputs are being achieved in accordance with the plan. The following aspects are analyzed in the process of monitoring of plans, policies, programmes and projects:

1. Whether or not resources are available to and used by the constituent units within the limits of an authorized budget and stipulated timeframe.
2. Whether or not expected outputs are achieved in a timely and cost-effective manner.
3. What is the level of implementation capacity?
4. What kind of problems and constraints are being faced and what kind of remedial measures are called for?

During monitoring, data and information on the above mentioned aspects are collected, processed and reported in a continuous, systematic, and time-bound manner. This helps identify problems and initiate corrective measures before it is too late.

### Evaluation

Evaluation is a systematic and purposeful undertaking carried out by internal or external evaluators to appraise the

- relevance, 
- efficiency,
- effectiveness of, as well as,
- impacts and sustainability

generated by the plans, policies, programmes and projects under implementation. The main objective of evaluation is to draw lessons from the strengths and weaknesses experienced in the implementation of plans, policies, programmes and projects so as to improve their design and implementation in the future as well as to hold the officials and agencies involved in the process accountable for its implementation and results.

```{r monitoring-evaluation-difference}

tribble(~"Monitoring", ~"Evaluation",
"It is a continuous process carried out throughout the implementation period and conceived as an internal function of the project.",
"Carried out at different stages such as ex-ante, on-going, terminal and ex-post.",
"It is performed only in the formative and operational stages.",
"Performed from selection stage and after the operation stage.",
"It is an institutional process to rectify gaps and weaknesses experienced during the formative and operational stages.",
"It is a means for initiating timely corrective measures during formative and operational stages as well as drawing lessons from the past strengths and weaknesses to inform future interventions.",
"It is directly related with inputs, processes and outputs.",
"It is concerned with the assessment of progress against pre-determined goals, outcomes and their effects and impacts on stakeholders.",
"It is a part of the management system.",
"It is an operational management tool to derive lessons from implemented plans, policies, programmes and projects.",
"It is generally conducted by the implementing agency.",
"It is generally conducted by independent individuals and institutions outside implementing agency."
) %>% 
 knitr::kable(booktabs = TRUE, caption = "Difference between monitoring and evaluation", longtable = TRUE) %>% 
 kableExtra::kable_styling()

```

Various types of monitoring and evaluations conducted at different stages of programme and project implementaiton is elaborated in \ref{tab:programme-project-implementaion-m-and-e}.

```{r programme-project-implementaion-m-and-e}
tribble(
~"Types of monitoring and evaluation", ~"Objectives",
"Continuous Monitoring", "To track performance against predetermined input, activity, process and output indicators during the formative phase of a programme/project.",
"Sustainability Monitoring", "To ensure necessary resources and provisions needed for sustainability throughout a programmes pre-determined lifecycle. It is conducted after the completion of the formative phase of programme/project.",
"Ex-ante Evaluation", "To determine the needs and assurances of programme/project continuity.\\newline To define indicators and clearly articulate the details of a given programme/project.",
"On-going Evaluation","To review progress of programme/project. \\newline To revisit and improve pre-determined action plan and/or make necessary changes in operational modalities.",
"Terminal Evaluation", "To carry out review of capacity, effectiveness, and sustainability of programme/project. \\newline To determine whether additional follow up is necessary after the completion of a programme/project.",
"Ex-post Evaluation", "To undertake review of impacts and sustainability of a programme/project. \\newline To obtain the lessons and recommendations for improvement of design and implementation of future interventions.",
"Continuous Feedback", "The feedback from different stages of monitoring and evaluation are to be addressed in due course"
)
```

### Development of monitoring and evaluation in Nepal

Programme/project monitoring and evaluation started with commencement of planned development in (2013/14 - 2017/18) but systematic efforts on the implementation of evaluation of development projects were only initiated in the Fifth Plan (1975-1980). During this period, the function of monitoring and evaluation was extended beyond NPC to ministries and agencies and a sytem of project monitoring based on expenditure weightage was in practice. After the Seventh Plan (1985-1990), the system of monitoring and evaluation based on pre-identified sectoral performance indicators was introduced.

Second phase of monitoring and evaluation system has been since eighth plan (1992 - 1997) with the objective to achieve the expected outcomes by putting in place a regular, effective, and efficient monitoring and evaluation mechanism from central to project levels and by streamlining implementation from the outset of project activities.

Third phase of monitoring and evaluation system places results achieved by program/projects on the focus. Numerous variations and revisions of this system are found, including

- Construction of policy and result matrices: 10th plan (2002-2007), and Three year interim plan (2007-2010)
- Poverty monitoring and analysis system (PMAS): 11th plan
- Technical audits of construction related projects by National Vigilance Centre (NVC)
- Financial Comptroller General Office (FCGO) introduced the Public Expenditure Tracking Survey (PETS) to analyze the status of public service delivery.
- NPC developed system of acquiring feedbacks

### Basis of Monitoring and Evaluation

Monitoring and evaluation of a given plan, policy, program and project can be carried out on the following bases. 

- Policy
 - While implementing policies of the state, it is necessary to monitor the various activities underway and to evaluate the outcomes and impacts produced by such activities and outputs.
- Periodic plan
 - Activities are carried out in various sectors to achieve the goals and objectives set out by the periodic plans. Regular monitoring of programme and project activities provides information on the extent of progress made towards achieving the goals and objectives of the plan.
 - Evaluation offers feedback on whether the programme and projects are adequately oriented towards achieving the goal of the plan
 - Mid-term review
- Business plan
 - Three year sectoral plans which are prepared in line with the assigned businesses or responsibilities of the concerned public agencies, as well as with goals and objectives set forth by the periodic plans
- Medium-term expenditure framework
 - Schema for prioritizing development programmes with a view to improve the process of programme/project formulation and implementation, improve the effectiveness of development programmes, and to ensure the budgets for the programmes and projects critical to the achievement of goals and objectives of periodic plan
- Budget policy and programme
 - Implementing ministries are required to report the progress of these policies and programmes to the NPC and the MOF at different intervals.
- Project document
 - Project documents elucidate goal, objectives, outputs, processes, and inputs of the project
 - Monitoring facilitates the timely completion of projects by keeping track of the activities and inputs, and by initiating timely corrective measures to resolve problems in implementation.
 - Evaluation helps in realizing time-relevant outputs from the projects through timely corrections as well as in obtaining feedback to inform other similar project designs
- Immediate action plan

### Current method and systems of Monitoring and Evaluation

- Monitoring and evaluation forms
 - MOF in its successive publications of the Budget Formulation Guidelines has unfailingly included Annual Development Programme (NPC ME form no. 1) and Progress Report Form (NPC ME form no. 2)
- Technical audit (by NVC)
- Public expenditure and tracking survey (by FCGO)
- Performance based budget released system
 - Since the beginning of Tenth Plan, the system of budget releases has been tied to project performance. For this purpose, performance indicators have been identified and arrangements have been made for performance-based budget releases so as to ensure that the set out objectives and targets in priority one (P1) projects are achieved within stipulated timeframes.
 - Provisions have been made to release the amount equivalent to one-third of the total authorized or the amount needed to carry out the targeted programmes/activities of the first trimester (whichever amount is greater) in the beginning of fiscal year. This is intended to ensure the availability of adequate financial resources for the respective office/projects. However, in case of subsequent trimesters, if the weighted progress of a given project is greater than 80 percent, the District Treasury and Controller Office (DTCO) automatically release the budget, if the weighted progress against the target is more than 50 percent and less than 80 percent, the DTCO will release budget only on the condition that the office or project requesting the budget receive authorization by letter from the parent ministry to that effect. If the weighted progress is below 50 percent of the set target, the concerned parent ministry needs to undertake an investigation and analysis of the reasons behind the failure to progress, and if necessary, take appropriate actions against the head of the office/project. Based on the investigation, if the ministry considers it appropriate to release the budget by ensuring mechanisms to accomplish tasks within the given timeline, it is required to inform the concerned DTCO by indicating the actual percentage of initial authorization that should be released. The DTCO will release the budget only when the office/project concerned makes a request by enclosing such a letter from the ministry.

### Results based management

- Underscores  the  need  to  enhance  the  effectiveness  of  investments  by linking outputs and outcomes with inputs and activities.
- Managing for Development Results (MfDR) is an important strategy of result based management, which constitutes following activities,
  - 

## Business planning

Simply stating, planning is the future road map of an organization. An organization, which runs on business principles, must be guided by its business planning. The planning which is outlined by using existing and available resources and information including vision, mission, strategy, objectives, goals or targets with income generating action plan of the organization is called business planning. It is expressed in term of long term planning or investment decision with budgeting for acquiring organization's future benefits over the years to come. The Term business planning or corporate planning was used in the 1960s, but now it has widely and popularly been referred to as strategic planning or strategic management. As a formal and structured approach, business planning or corporate planning or strategic planning can be instrumental in achieving strategic corporate goals and objectives either for any service oriented public organization or for any profit oriented business organization. 

Business planning focuses on efficiently and effectively achieving optimal performance from all business oriented divisions or units of organization with actionable information provided to executive management to foster accountability and business growth. Most effective and actionable plans are developed by gathering various information with addressing organization's goals and objectives usually by the team of executives or senior management or by the management experts. Various statistical data including budget formulation and performance management is used in connection with analysis and forecasting for business planning.

The business plan and budget must be co-joint to each other and accordingly the framework of business strategy should be developed for short term and long term business plan. The project to be carried out during the specified period should be indentified within the limitation of budgetary framework.

The process of drawing up detailed action plans to achieve an organization's goals and objectives, taking into account the resources of the organization and the environment within which it operates, are minutely scrutinized for coping with the organization's future requirements. So business planning represents a formal, structured approach to achieving objectives and to implementing the corporate strategy for the long term benefit of the organization.


An effective business planning mainly includes the following aspects of the organization:
- Forecasting the strategic vision of the organization.
- Laying out the communication Tools of the organization.
- Setting short term and long term planning by involving to all levels of management staff.
- Finding the sources of funds for phase wise investment.
- Managing the various resources of the organization
- Preparing budget for sound implementation of such planned projects.
- Forecasting and reviewing projects as per the need of the organization
- Enhancing all employees for its effective implementation
- Finding the basis for rational decision making process.

The organization may either be small or big in size, but business planning is the first and foremost as well as essential step or tool for gradual development of the organization. So it should be judged highly by the top level management. It may be simplified by enriching special format and may also be sub-classified as per the requirement of the particular directorate, departments, divisions and various grassroots offices as well. 

The business planning may have several purposes. It encourages the management team to focus on why they are in business or in commercial activities and what it takes to succeed. It is the vehicle which describes product or service or technology to others and why they will buy it.? It also provides the guidelines to management on how an organization will managerially and technologically be managed. Similarly it describes the financial plans of the organization and also convinces investors to invest huge amount of fund in the organization. 

It serves as a guideline for allocating resources in a coordinated way. It handles unforeseen complications, helps make appropriate business decisions and provides specific and organized information about the organization how investors will be rewarded. It specifies ends and means of implementing actions. Business planning strategically defines the existing state and desired state of the organization. So, an appropriate business planning is a crucial part of the organization. 

Now-a-days, business planning is interchangeably used as strategic planning, but its literal meaning is slightly different. Strategic planning is an organization's process of defining its strategy, or direction, and making decisions on allocating its resources to achieve its final goal, including its financial resources or capital and human resources or people in a long term perspective. Various business analysis techniques can be used in strategic planning, including SWOT analysis (Strengths, Weaknesses, Opportunities, and Threats) and PEST analysis (Political, Economic, Social, and Technological analysis) or STEER analysis involving Socio-cultural, Technological, Economic, Ecological, and Regulatory factors. Generally strategic planning adopts the following process in its initial formulation stage.

A good business planning may address many issues such as innovation assessment, market research, competitive analysis, business model, financial projections and management team. Business planning may be focused externally or internally as stated above. Externally focused plans target goals that are important to external financial stakeholders. They typically have detailed information about the organization or team attempting to reach the goals. For profit earning entity, external stakeholders include investors and customers. External stakeholders of non-profit making entity include donors and the clients of the non-profit services. For government agencies, external stakeholders include tax-payers, higher-level government agencies, and international lending bodies such as Asian Development Bank, World Bank etc.

Internally focused business plans target intermediate goals required to reach the external goals. They may cover the development of a new product, a new service, a new information technology system, a restructuring of financing methods or a reengineering of the organization structure. An internal business planning is often developed in conjunction with a balanced scorecard. This allows success of the plan to be measured using non-financial measures. Business planning identifies and targets internal goals, but provides only general guidance on how they will be fulfilled which are called strategic planning.

To sum up, business planning is one of the most crucial decision-making tools. It represents all aspects of business planning process; declaring vision, mission, values and strategy alongside sub-plans to cover marketing, finance, operation, regulation, human resources, infrastructure and legal plan. Business planning draws wide range of knowledge from many different disciplines such as finance, human resource management, economics, law, corporate, engineering, intellectuals, stakeholders, regulators, service providers, operation and marketing management etc for its preparation. A good business planning can help to make a good business credible, understandable, and attractive for enhancing or fostering commercial activities of the organization. A business planning is a macro planning aspect of all disciplinary plans. It covers wide field. This article is just a theory based pictorial aspect of business planning.
